The College Application Problem

### 2022-05-06 slides_20min/slides.tex
script, and arXiv URL
typos; garbage collect unused latex imports

\ifen 
\title{The College Application Problem}
\author{Max Kapur}
\institute{Seoul National University\\
Department of Industrial Engineering \\
Management Science/Optimization Lab\\
\url{maxkapur@snu.ac.kr}}
\date{\today}
\else
\title[대학 지원 최적화 문제]{대학 지원 최적화 문제}
\author{Max Kapur}
\institute{서울대학교 산업공학과 \\
경영과학/최적화 연구실\\
\url{maxkapur@snu.ac.kr}}
\date{\today}
\fi

@@

 \setbeamertemplate{caption}[numbered]
\setbeamertemplate{footline} {
  \leavevmode%
  \hbox{%
  \begin{beamercolorbox}[wd=.25\paperwidth,ht=2.65ex,dp=1.2ex,left]{author in head/foot}%
  \hspace*{1.5ex}\insertshortauthor
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.27\paperwidth,ht=2.65ex,dp=1.2ex,left]{date in head/foot}%
  \hspace*{1ex}
    \ifnum \theframenumber=1{}
    \else \insertshorttitle%~\textbar~\insertsectionhead
    \fi
     \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.23\paperwidth,ht=2.65ex,dp=1.2ex,right]{date in head/foot}%
    \ifnum \theframenumber=1{}
    \else \insertsectionhead
    \fi
  \hspace*{2ex}
     \end{beamercolorbox}%
       \begin{beamercolorbox}[wd=.15\paperwidth,ht=2.65ex,dp=1.2ex,left]{author in head/foot}%
  \hspace*{1.5ex}\insertdate
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.1\paperwidth,ht=2.65ex,dp=1.2ex,right]{author in head/foot}%
    \insertframenumber{} / \inserttotalframenumber
    \hspace*{1.5ex}
  \end{beamercolorbox}
}%
  \vskip0pt%
}

@@

\ifen {
\begin{frame}[plain]
  \vfill
\begin{center}
~
\begin{beamercolorbox}[wd=.5\textwidth,sep=8pt,center,shadow=false,rounded=true]{section in head/foot}
    \usebeamerfont{title} \insertsectionhead
  \end{beamercolorbox}
~
\end{center}
\end{frame}
} \else {
\begin{frame}[plain]
  \vfill
\begin{center}
~
\begin{beamercolorbox}[wd=.5\textwidth,sep=8pt,center,shadow=false,rounded=true]{section in head/foot}
    \usebeamerfont{title} \insertsectionhead
  \end{beamercolorbox}
~
\end{center}
\end{frame}
} \fi

@@

The student's overall utility is the $t_j$-value associated with the \textbf{best} school she gets into:
\[\text{Utility} =\max\bigr\{t_0, \max\{t_j Z_j : j \in \mathcal{X}\}\bigr\}\]

When the student's application portfolio is $\mathcal{X}$, we refer to her expected utility as the portfolio's \textbf{valuation} $v(\mathcal{X})$. 
\end{frame}
} \else {
\begin{frame}{효용 모형}
$c_j$에 다니면 $t_j \geq 0$ 단위의 \textbf{효용}이 발생한다. Wlog, $t_j \leq t_{j+1}$.

@@

\ifen {
\begin{frame}[plain]
  \vfill
\begin{center}
~
\begin{beamercolorbox}[wd=.5\textwidth,sep=8pt,center,shadow=false,rounded=true]{section in head/foot}
    \usebeamerfont{title} \insertsectionhead
  \end{beamercolorbox}
\end{center}
\end{frame}
} \else {
\begin{frame}[plain]
  \vfill
\begin{center}
~
\begin{beamercolorbox}[wd=.5\textwidth,sep=8pt,center,shadow=false,rounded=true]{section in head/foot}
    \usebeamerfont{title} \insertsectionhead
  \end{beamercolorbox}
\end{center}
\end{frame}
} \fi

\begin{frame}[plain]%{Safety, target, and reach schools}
\vfill

@@

\ifen {
\begin{frame}[plain]
  \vfill
\begin{center}
~
\begin{beamercolorbox}[wd=.5\textwidth,sep=8pt,center,shadow=false,rounded=true]{section in head/foot}
    \usebeamerfont{title} \insertsectionhead
  \end{beamercolorbox}
\end{center}
\end{frame}
} \else {
\begin{frame}[plain]
  \vfill
\begin{center}
~
\begin{beamercolorbox}[wd=.5\textwidth,sep=8pt,center,shadow=false,rounded=true]{section in head/foot}
    \usebeamerfont{title} \insertsectionhead
  \end{beamercolorbox}
\end{center}
\end{frame}
} \fi

@@

Using a variable-elimination technique, we reduce the cost of computing $v(\mathcal{X})$ to amortized unit time, obtaining an $O(hm)$ algorithm. Our Julia implementation (Kapur 2022) solves an $m = 16384$ instance in 200 ms.

@@

\ifen {
\begin{frame}[plain]
  \vfill
\begin{center}
~
\begin{beamercolorbox}[wd=.5\textwidth,sep=8pt,center,shadow=false,rounded=true]{section in head/foot}
    \usebeamerfont{title} \insertsectionhead
  \end{beamercolorbox}
\end{center}
\end{frame}
} \else {
\begin{frame}[plain]
  \vfill
\begin{center}
~
\begin{beamercolorbox}[wd=.5\textwidth,sep=8pt,center,shadow=false,rounded=true]{section in head/foot}
    \usebeamerfont{title} \insertsectionhead
  \end{beamercolorbox}
\end{center}
\end{frame}
} \fi

@@

Kapur, Max and Sung-Pil Hong. 2022. ``The Optimal College Application Problem.'' Preprint. \url{https://github.com/maxkapur/CollegeApplication}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{frame}{\ifen Introduction\else 서론 \fi}
\ifen {
The optimal college application problem is a \textbf{novel combinatorial optimization problem}.

Our methodological orientation:
\begin{itemize}
\item \textbf{Asset allocation perspective:} Valuation vs. value, inherent risk/reward tradeoff, efficient frontier.
\item \textbf{Knapsack perspective:} Integrality constraint, dynamic programming solutions, NP-completeness, approximation algorithms.
\end{itemize}
} \else {

방법론적 지향:
\begin{itemize}
\item \textbf{자산배분 최적화 관점:} 가치 (valuation) vs. 가치 (value), 위험/보상 관리, 효율적 투자선.
\item \textbf{배낭 문제 관점:} 정수 조건, 동적 계획, NP-completeness, 근사 해법.
\end{itemize}
} \fi
\end{frame}

\begin{frame}{\ifen The college application process\else 대학 지원 과정\fi}
\ifen Consider a student deciding where to apply to college. 
\else 대학 지원 전략을 고민하는 학생을 고려하자. \fi

\ifen There are $m$ schools, $\mathcal{C} = \{ 1 \dots m\}$. The $j$th school is named $c_j$.
\else $m$개의 학교, $\mathcal{C} = \{ 1 \dots m\}$. $j$번째 학교의 이름은 $c_j$. \fi

\ifen $t_j > 0$: \textbf{Utility} from attending $c_j$. Receive $t_0$ units of utility if she gets in nowhere.  Wlog, $t_0 < t_1 \leq \cdots \leq t_m$. 
\else 학생이 $c_j$에 다니면 그의 효용은 $t_j > 0$이며 대학에 못가면 $t_0$. Wlog, $t_0 < t_1 \leq \cdots \leq t_m$. \fi

\ifen $f_j \in (0, 1]$: \textbf{Probability of getting into} $c_j$ if she applies. $f_0 = 1$. Admissions outcome is $Z_j \sim \operatorname{Bernoulli}(f_j)$. 
\else $f_j \in (0, 1]$: 학생이 $c_j$에 지원할 때, 그의 \textbf{합격 확률}. $f_0 = 1$. 합격 결과는 $Z_j \sim \operatorname{Bernoulli}(f_j)$. \fi

\ifen $\mathcal{X}$: Student's \textbf{application portfolio}, the set of schools to which she chooses to apply. 
\else $\mathcal{X}$: 학생의 \textbf{지원 포트폴리오}, 즉 지원하는 학교의 집합.\fi

\ifen $p_j(\mathcal{X})$: \textbf{Probability that student attends} $c_j$. For $j= 0 \dots m$,
\else $p_j(\mathcal{X})$: 학생이 $c_j$에 \textbf{진학할 확률}. $j= 0 \dots m$에 대해,\fi
\begin{align*}
p_j(\mathcal{X}) &= 
\begin{cases}
\displaystyle f_j  \prod_{\substack{j’ \in \mathcal{X}: \\ j' > j}} (1 - f_{j’}), \quad & j \in \{0\}\cup\mathcal{X}\\
0, \quad & \text{otherwise.}
\end{cases} 
\end{align*}
\end{frame}

\begin{frame}{\ifen Objective function and constraints\else 목적함수와 제약 조건\fi}
\ifen {
The student's objective is to maximize her expected utility, which is the expected value of the \textbf{best} school she gets into. Notice the ``maximax'' form.
\begin{align*}
v(\mathcal{X})
&= \operatorname{E}\Bigl[\max\bigr\{t_0, \max\{t_j Z_j : j \in \mathcal{X}\}\bigr\}\Bigr] \\
&= \sum_{j=0}^m t_j p_j(\mathcal{X}) = \sum_{j\in\{0\}\cup\mathcal{X}} \Bigl( f_j t_j \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)  \label{closedformportfoliovaluationX}%, \quad \text{or equivalently,}\\
%\qquad v(x) &= t_0 \prod_{j=1}^m (1 - f_{j} x_j) + \sum_{j=1}^m \Bigl( x_j t_j f_j \prod_{j’ = j+1}^m (1 - f_{j’} x_{j’}) \Bigr) \label{closedformportfoliovaluationx}
\end{align*}
(As we show, wlog to assume $t_0 = 0$.)

We consider two kinds of constraints:
\begin{itemize}
\item \textbf{Cardinality constraint}: A limit $h$ on the number of applications, as in Korea's college admissions process. ``Alma's problem,'' solvable in polynomial time.
\item Application \textbf{budget constraint}: Each school has application cost $g_j$, and student has $H$ units to spend on applications. ``Ellis's problem,'' NP-complete.
\end{itemize}
}\else{
학생의 목적은 자신이 기대 효용을 최대화하는 것이며 이는 합격하는 학교 중 효용이 \textbf{최대}인 학교이다. 목적함수의 ``maximax'' 형태는 본 모형의 특색.
\begin{align*}
v(\mathcal{X})
&= \operatorname{E}\Bigl[\max\bigr\{t_0, \max\{t_j Z_j : j \in \mathcal{X}\}\bigr\}\Bigr] \\
&= \sum_{j=0}^m t_j p_j(\mathcal{X}) = \sum_{j\in\{0\}\cup\mathcal{X}} \Bigl( f_j t_j \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)  \label{closedformportfoliovaluationX}%, \quad \text{or equivalently,}\\
%\qquad v(x) &= t_0 \prod_{j=1}^m (1 - f_{j} x_j) + \sum_{j=1}^m \Bigl( x_j t_j f_j \prod_{j’ = j+1}^m (1 - f_{j’} x_{j’}) \Bigr) \label{closedformportfoliovaluationx}
\end{align*}
(논문에서 보이듯이, $t_0 = 0$이라고 가정할 수 있다.)

2가지의 제약 조건 형태를 고려한다.
\begin{itemize}
\item \textbf{집합 크기 제약}: 한국 정시 입학 과정처럼 지원할 수 있는 학교의 개수가 $h$로 제한된 ``알마의 문제.'' 다항 시간 해법 존재.
\item \textbf{지원 비용 예산 조건}: 각 학교의 지원 비용이 $g_j$이며 학교가 지원에 쓸 수 있는 금액이 $H$인 ``엘리스의 문제.'' NP-complete.
\end{itemize}
}\fi
\end{frame}

\begin{frame}{\ifen Problem statement\else 문제 정의 \fi}
\begin{problem}[\ifen Alma’s problem\else 알마의 문제\fi]
\vspace{-1em}
%\ifen Alma's optimal college application portfolio is given by the solution to the following combinatorial optimization problem:
%\else
%알마의 최적 대학 지원 포트폴리오는 다음 조합 최적화 문제의 최적해이다.
%\fi
\begin{align*}
\begin{split}
\text{maximize}\quad &  v(\mathcal{X}) = \sum_{j\in
%\{0\}\cup
\mathcal{X}} \Bigl( f_j t_j \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)\\
\text{subject to}\quad & \mathcal{X}\subseteq\mathcal{C}, \quad|\mathcal{X}| \leq h 
\end{split}
\end{align*}
\end{problem}
\begin{problem}[\ifen Ellis's problem\else 엘리스의 문제\fi]
\vspace{-1em}
%\ifen
%Ellis's optimal college application portfolio is given by the solution to the following combinatorial optimization problem.
%\else 
%엘리스의 최적 대학 지원 포트폴리오는 다음 조합 최적화 문제의 최적해이다.
%\fi
\begin{align*}
\begin{split}
\text{maximize}\quad &  v(\mathcal{X}) = \sum_{j\in
%\{0\}\cup
\mathcal{X}} \Bigl(f_j t_j \prod_{\substack{i \in \mathcal{X}: \\ i > j}} (1 - f_{i}) \Bigr)\\
\text{subject to}\quad & \mathcal{X}\subseteq\mathcal{C},~~\sum_{j \in \mathcal{X}} g_j \leq H 
\end{split}
\end{align*}
\end{problem}

\ifen Set all $g_j = 1$ to see that Alma's problem is a special case of Ellis's problem. 
\else 모든 $g_j = 1$로 넣으면 알마의 문제가 엘리스의 문제의 특수한 경우임을 알 수 있다.  \fi
\end{frame}

\begin{frame}{\ifen Why college application is difficult\else 대학 지원 문제가 왜 어려운가\fi}
\ifen
``Obvious'' algorithm which picks the schools having highest expected value $f_j t_j$ is \textbf{incorrect.} ($\tfrac{1}{h}$-opt. for Alma’s problem; $\tfrac{1}{m-1}$-opt. for Ellis’s problem.)

College application can be expressed as a nonlinear integer program, but it is \textbf{not concave}.

\textrightarrow~Problem is \textbf{combinatorial} in nature.

Typical instances (see following slide) involve \textbf{negative correlation} between probabilities $f_j$ and utility values $t_j$.

\textrightarrow~Students must trade off attractive, selective “reach schools” against less preferable “safety schools” where admission is a safer bet.

A good application strategy has \textbf{monetary value}. US admissions consultants charge an average of \$200/hr (Sklarow 2018).

Previous research considered small ($m=8$) instances and solved by enumeration (Chao 2014).
\else
기대 가치 $f_j t_j$가 가장 큰 학교를 선택하는 `뻔한' 알고리즘은 사실 틀리다. (근사 계수는 알마 문제에 대해 $\tfrac{1}{h}$, 엘리스 문제에 대해 $\tfrac{1}{m-1}$.)

정수 비선형 계획으로 표현할 수 있지만 목적함수가 \textbf{오목}이 아니다.

\textrightarrow~본성적으로 \textbf{조합적인} 문제.

%\textbf{알마의 문제는 다항 시간} 안에 풀 수 있지만, 기대 가치 $f_j t_j$가 가장 큰 $h$개의 학교를 선택하는 `뻔한' 알고리즘은 사실 틀리다.

%\textbf{엘리스의 문제는 NP-complete}하며 배낭 문제에서 변환함으로 증명.

전형적인 인스턴스(다음 화면)에서 입학 확률 $f_j$와 효용 모수 $t_j$가 서로 \textbf{반비례}한다.

\textrightarrow~선호도가 높으며 붙기 어려운 “상향” 지원 학교(reach school)와 선호도가 낮으며 붙기 쉬운 “안정” 지원 학교(safety school) 사이의 균형을 고려해야 한다고 의미.

좋은 지원 전략은 \textbf{금전적인 가치}를 가진다. 미국 입학 상담가의 시간당 급료는 평균 24만원 (Sklarow 2018).

선행 연구는 작은 ($m=8$) 인스턴스만 고려하고 열거법으로 풀었다 (Chao 2014).
\fi
\end{frame}

% Reverse order of slides in Korean version
\ifen 
\begin{frame}[plain]{}
\begin{center}
\includegraphics[width=\textwidth]{plots/news-en.png}
\end{center}
\end{frame}
\fi

\begin{frame}[plain]{}
\begin{center}
\includegraphics[width=\textwidth]{plots/news-ko.png}
\end{center}
\end{frame}

\ifen \else
\begin{frame}[plain]{}
\begin{center}
\includegraphics[width=\textwidth]{plots/news-en.png}
\end{center}
\end{frame}
\fi

\ifen \section{Alma's problem} \else \section{알마의 문제} \fi

\ifen{
\begin{frame}[plain]
  \vspace{9em}
\begin{center}
~
\begin{beamercolorbox}[wd=.6\textwidth,sep=8pt,center,shadow=false,rounded=true]{section in head/foot}
    \usebeamerfont{title} \insertsectionhead \par%
  \end{beamercolorbox}
~
\end{center}
\end{frame}
}
\else{
\begin{frame}
  \vspace{9em}
\begin{center}
~
\begin{beamercolorbox}[wd=.6\textwidth,sep=8pt,center,shadow=false,rounded=true]{section in head/foot}
    \usebeamerfont{title} \insertsectionhead \par%
  \end{beamercolorbox}
~
\end{center}
\end{frame}
}
\fi

\begin{frame}{\ifen The nestedness property for Alma's problem\else 알마 문제 최적해의 포함 사슬 관계\fi}
\ifen The following theorem says that the solution to Alma's problem is \textbf{nested} in the application limit $h$.
\else 다음 정리는 알마 문제의 최적해가 지원 제한 $h$로 모수화된 \textbf{포함 사슬 관계}를 가진다고 의미한다. \fi
\begin{theorem}[\ifen Nestedness of optimal application portfolios\else 최적 포트폴리오의 포함 사슬 관계\fi] \label{nestedapplication}
\ifen There exists a sequence of portfolios $\{\mathcal{X}_h\}_{h=1}^m$ satisfying the nestedness relation
\else  각 $\mathcal{X}_h$가 지원 제한 $h$에 대한 최적 포트폴리오며 위의 포함 사슬 관계를 만족하는 포트폴리오 수열  $\{\mathcal{X}_h\}_{h=1}^m$가 존재한다.\fi
\begin{equation*}
\mathcal{X}_1 \subset \mathcal{X}_2\subset \dots \subset \mathcal{X}_m
\end{equation*}
\ifen such that each $\mathcal{X}_h$ is an optimal application portfolio when the application limit is $h$.\fi
\end{theorem}

\ifen
This implies the validity of a greedy algorithm that adds schools one at a time, maximizing $v(\mathcal{X})$ at each addition.

Some additional refinement reduces the computation time to $O(hm)$.
\else
따라서 $v(\mathcal{X})$를 최대화하는 학교를 차례대로 추가하는 탐욕 해법의 타당성이 성립.

이를 좀 더 조정하면 계산 시간을 $O(hm)$으로 절감할 수 있다.
\fi
\end{frame}

\begin{frame}{\ifen Concavity of $v(\mathcal{X}_h)$\else $v(\mathcal{X}_h)$의 오목성\fi}
\ifen The nestedness property implies that Alma's expected utility is a discretely concave function of $h$.
\else 포함 사슬 관계 성질은 알마의 기대 효용이 $h$의 이산 오목 함수임을 의미한다.\fi

\begin{theorem}[\ifen Optimal portfolio valuation concave in $h$\else 최적 포트폴리오 가치의 $h$-오목성\fi] \label{concavityinh}
\ifen For $h = 2 \dots (m-1)$,
\else $h = 2 \dots (m-1)$에 대해,\fi
\begin{equation*}v(\mathcal{X}_h) - v(\mathcal{X}_{h-1}) \geq v(\mathcal{X}_{h+1}) - v(\mathcal{X}_{h}).\end{equation*} 
\end{theorem}

\ifen
Corollary: $v(\mathcal{X}_h)$ is $O(h)$. Paper contains a tight example. 
\else
이의 따름정리로서 $v(\mathcal{X}_h)$가 $O(h)$ 함수가 되며, 논문에서 타이트한 예를 제시. 
\fi
\end{frame}

\begin{frame}{A small example}
\ifen
College data and optimal application portfolios for a fictional market with $m=8$ schools.
\else
$m=8$개의 학교로 이루어진 가상 입학 시장의 대학교 자료와 최적 지원 포트폴리오.
\fi
\begin{center}
\begin{tabular}{r|lcccc}
\ifen\textbf{Index $j$} & \textbf{School $c_j$} & \textbf{Admit prob. $f_j$} & \textbf{Utility $t_j$} & \textbf{Priority} & \textbf{$v(\mathcal{X}_h)$} \\ \hline
\else 
\textbf{지표 $j$} & \textbf{학교 $c_j$} & \textbf{합격 확률 $f_j$} & \textbf{효용 $t_j$}  & \textbf{지원 순위} & \textbf{$v(\mathcal{X}_h)$} \\ \hline  \fi
\\[-.75em]
1 & \ifen Mercury University   \else  수성대  \fi   & 0.39   & 200 & 4   & 230.0   \\
2 & \ifen Venus University     \else  금성대  \fi   & 0.33   & 250 & 2   & 146.7  \\
3 & \ifen Mars University      \else  화성대  \fi   & 0.24   & 300 & 6   & 281.5  \\
4 & \ifen Jupiter University   \else  목성대  \fi   & 0.24   & 350 & 1   & \phantom{0}84.0  \\
5 & \ifen Saturn University    \else  토성대  \fi   & 0.05   & 400 & 7   & 288.8  \\
6 & \ifen Uranus University    \else  천왕성대 \fi   & 0.03   & 450 & 8   & 294.1  \\
7 & \ifen Neptune University   \else  해왕성대 \fi   & 0.10   & 500 & 5   & 257.7  \\
8 & \ifen Pluto College        \else  명왕성대 \fi   & 0.12   & 550 & 3   & 195.1      
\end{tabular}
\end{center}

\ifen 
By the nestedness property, the optimal portfolio when the application limit is $h$ consists of the $h$ schools having priority $h$ or less.

The following slide illustrates the concavity property.
\else
포함 사슬 관계 성질에 따라, 지원 제한이 $h$일 때 최적 포트폴리오는 지원 순위가 $h$ 이하인 $h$개의 학교로 구성된다.

다음 화면에서 나타나는 곡선은 오목성 성질을 보인다.
\fi
\end{frame}

\section{Ellis's problem}

\ifen{
\begin{frame}[plain]
  \vspace{9em}
\begin{center}
~
\begin{beamercolorbox}[wd=.6\textwidth,sep=8pt,center,shadow=false,rounded=true]{section in head/foot}
    \usebeamerfont{title} \insertsectionhead \par%
  \end{beamercolorbox}
~
\end{center}
\end{frame}
}
\else{
\begin{frame}[plain]
  \vspace{9em}
\begin{center}
~
\begin{beamercolorbox}[wd=.6\textwidth,sep=8pt,center,shadow=false,rounded=true]{section in head/foot}
    \usebeamerfont{title} \insertsectionhead \par%
  \end{beamercolorbox}
~
\end{center}
\end{frame}
}
\fi

\begin{frame}{\ifen Three algorithms for Ellis's problem \else 엘리스의 문제를 위한 3개의 알고리즘\fi}
\ifen {
By setting $f_j = \varepsilon$, we can make $v(\mathcal{X})$ arbitrarily close to a linear function. This means that knapsack reduces to Ellis's problem, and therefore Ellis's problem is NP-complete.

We provide three algorithms for Ellis's problem.
\begin{itemize}
\item \textbf{Branch-and-bound} routine. Theoretically interesting but not very efficient.
\item \textbf{Dynamic program} based on total expeditures. Pseudopolynomial time, namely $O(Hm + m\log m)$, and exact. Very efficient for typical instances in which $g_j$ are small integers.
\item Another dynamic program based on truncated portfolio valuations. Yields a $(1 - \varepsilon)$-optimal solution in $O(m^3 /\varepsilon)$-time: \textbf{an FPTAS}.
\end{itemize}
} \else {
$f_j = \varepsilon$으로 넣으면 $v(\mathcal{X})$를 선형 함수에 원하는 만큼 가깝게 만들 수 있다. 배낭 문제가 엘리스의 문제로 변환할 수 있음을 의미하므로 엘리스의 문제는 NP-complete하다.

엘리스 문제를 위해 3가지해법을 제시한다.
\begin{itemize}
\item \textbf{분지한계법}은 이론적으로 흥미롭지만 현실적 효율성이 낮다.
\item 지원 지출액 기반 \textbf{동적 계획}. $O(Hm + m\log m)$인 의사 다항 시간 안에 정확한 해를 출력하며, $g_j$가 작은 정수가 되는 전형적인 인스턴스에 매우 효율적이다.
\item 분수값을 내림한 포트폴리오 가치 기반 다른 동적 계획. $O(m^3 /\varepsilon)$-시간 안에  $(1 - \varepsilon)$-근사 해를 출력하므로 \textbf{FPTAS}.
\end{itemize}

} \fi
\end{frame}

\ifen \section{Ellis's problem: Branch and bound} \else \section{엘리스의 문제: 분지한계법} \fi

\begin{frame}{\ifen Branch-and-bound\else 분지한계법\fi}
\ifen
Our branch-and-bound framework is based on the following integer formulation of Ellis's problem. 
\else
엘리스의 문제는 다음 정수 비선형 계획으로 표현되며 이를 바탕으로 분지한계법을 만든다. 
\fi
\begin{problem}[\ifen Integer NLP for Ellis's problem\else 엘리스의 문제를 위한 비선형 정수 계획\fi]% \label{integernlp}
\vspace{-1em}
\begin{align*}
\begin{split}
\text{maximize}\quad &  v(x) = \sum_{j=1}^m \Bigl(f_j t_j  x_j \prod_{i > j} (1 - f_{i} x_i) \Bigr)\\
\text{subject to}\quad & \sum_{j=1}^m g_j x_j \leq H, ~~ x_i \in \{0, 1\}^m
\end{split}
\end{align*}
\end{problem}

\ifen
Objective is a nonconvex polynomial of degree $m$ \dots~numerical solvers are hopeless!
\else
목적함수는 $m$차 비오목 다항식 \dots~수리적 솔버는 헛수고!
\fi
\end{frame}

\begin{frame}{\ifen LP relaxation \else 선형 완화 문제\fi}
\ifen 
We use the following LP upper bound.
\else
다음 선형 완화 문제를 활용한다.
\fi
\begin{problem}[\ifen LP relaxation for Ellis's problem\else 엘리스의 문제를 위한 선형 완화 문제\fi] %\label{LPrelaxation}
\vspace{-1em}
\begin{align*}
\begin{split}
\text{maximize}\quad &  v_{\mathrm{LP}}(x) = \sum_{j=1}^m  f_j t_j x_j \\
\text{subject to}\quad & \sum_{j=1}^m g_j x_j \leq H, ~~ x \in [0, 1]^m
\end{split}
\end{align*}
\end{problem}
\ifen
This is a continuous knapsack problem, easy to solve (Dantzig 1957; Balas and Zemel 1980).

We \textbf{tighten this bound} by recycling a variable-elimination technique from the proof of the nestedness theorem.

A straightforward implementation is OK on small instances ($m \leq 35$). \textbf{Considerable room for improvement} by using better heuristics to select the branch node.
\else
연속 배낭 문제라고 불리며 쉽게 풀 수 있다 (Dantzig 1957; Balas and Zemel 1980).

포함 사슬 관계의 증명 과정에서 도출한 변수 소거법은 재활용해서 위에 \textbf{상한을 더 타이트하게} 조정한다.

이 기반으로 만든 간단한 알고리즘은 작은 ($m \leq 35$) 인스턴스에 괜찮지만 분지 마디를 선택하는 휴리스틱으로 \textbf{개선할 여지}가 보인다.
\fi
\end{frame}

\ifen \section{Ellis's problem: Expenditures DP} \else \section{엘리스의 문제: 지원 지출액 동적 계획}\fi

\begin{frame}{\ifen Expenditures-based dynamic program\else 지원 지출액 기반 동적 계획\fi}
\ifen Let $V[j,h] $ denote the value of the optimal portfolio that uses only the schools $\{ 1, \dots, j\}$ and costs no more than $h$.
\else $\{ 1, \dots, j\}$에 속한 학교만 사용하면서 지원 지출액이 $h$를 넘지 않은 최적 포트폴리오의 가치를 $V[j, h]$라고 하자.\fi

\ifen Then we can use the following \textbf{Bellman equation} to compute all the $V[j, h]$-values recursively.
\else 그러면 다음과 같은 \textbf{Bellman 식}으로 모든 $V[j, h]$-값은 재귀적으로 계산할 수 있다. \fi
\begin{align*}
V[j, h] = \max\bigl\{ V[j-1, h], (1 - f_j) V[j-1, h-g_j] + f_j t_j \bigr\}
\end{align*}
\ifen
This equation works because the $t_j$-values are indexed in ascending order.

Filling a lookup table with the $V[j, h]$-values therefore costs $O(Hm + m\log m)$, and then computing $\mathcal{X}$ is trivial.

\textbf{Very effective} on typical instances where $g_j$ are small integers.
\else
이 식의 타당성은 학교를 $t_j$-값 순서대로 배열함에 의존.

따라서 $V[j, h]$-값으로 표를 채우는 시간이 $O(Hm + m\log m)$이며 이를 참고하면 $\mathcal{X}$를 쉽게 구할 수 있다.

전형적인 인스턴스에서 $g_j$가 작은 상수이므로 \textbf{매우 효율적인} 해법.
\fi
\end{frame}

\ifen \section{Ellis's problem: Valuations DP} \else \section{엘리스의 문제: 포트폴리오 가치 기반 동적 계획}\fi

\begin{frame}{\ifen Valuation-based dynamic program \else 포트폴리오 가치 기반 동적 계획 \fi}
\ifen 
As with the knapsack problem, Ellis's problem admits a \textbf{complementary dynamic program} that iterates on the value of the cheapest portfolio instead of on the cost of the most valuable portfolio.

>>>>

\ifen 
\title{The College Application Problem}
\author{Max Kapur and Sung-Pil Hong}
\institute{Seoul National University\\
Department of Industrial Engineering \\
Management Science/Optimization Lab\\
\url{maxkapur@snu.ac.kr}}
\date{\today\\
Spring Conference of the Korean Institute of Industrial Engineers}
\else
\title[대학 지원 최적화 문제]{대학 지원 최적화 문제\\The College Application Problem}
\author{Max Kapur, 홍성필}
\institute{서울대학교 산업공학과\\
경영과학/최적화 연구실\\
\url{maxkapur@snu.ac.kr}}
\date{\today\\
대한산업공학회 춘계공동학술대회}
\fi

@@

 \setbeamertemplate{caption}[numbered]
\setbeamertemplate{footline} {
  \leavevmode%
  \hbox{%
  \begin{beamercolorbox}[wd=.25\paperwidth,ht=2.65ex,dp=1.2ex,left]{author in head/foot}%
  \hspace*{1.5ex}\insertshortauthor
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.27\paperwidth,ht=2.65ex,dp=1.2ex,left]{date in head/foot}%
  \hspace*{1ex}
    \ifnum \theframenumber=1{}
    \else \insertshorttitle%~\textbar~\insertsectionhead
    \fi
     \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.23\paperwidth,ht=2.65ex,dp=1.2ex,right]{date in head/foot}%
    \ifnum \theframenumber=1{}
    \else \insertsectionhead
    \fi
  \hspace*{2ex}
     \end{beamercolorbox}%
       \begin{beamercolorbox}[wd=.15\paperwidth,ht=2.65ex,dp=1.2ex,left]{author in head/foot}%
  \hspace*{1.5ex}%\insertdate
  \end{beamercolorbox}%
  \begin{beamercolorbox}[wd=.1\paperwidth,ht=2.65ex,dp=1.2ex,right]{author in head/foot}%
    \insertframenumber{} / \inserttotalframenumber
    \hspace*{1.5ex}
  \end{beamercolorbox}
}%
  \vskip0pt%
}

@@

\ifen {
\begin{frame}[plain]
\begin{center}
~
\begin{beamercolorbox}[wd=.5\textwidth,sep=8pt,center,shadow=false,rounded=true]{section in head/foot}
    \usebeamerfont{title} \insertsectionhead
  \end{beamercolorbox}
~
\end{center}
\end{frame}
} \else {
\begin{frame}[plain]
\begin{center}
~
\begin{beamercolorbox}[wd=.5\textwidth,sep=8pt,center,shadow=false,rounded=true]{section in head/foot}
    \usebeamerfont{title} \insertsectionhead
  \end{beamercolorbox}
~
\end{center}
\end{frame}
} \fi

@@

The student's overall utility is the $t_j$-value associated with the \textbf{best} school she gets into:
\[\text{Utility} =\max\bigr\{t_0, \max\{t_j Z_j : j \in \mathcal{X}\}\bigr\}\]
When the student's application portfolio is $\mathcal{X}$, we refer to her expected utility as the portfolio's \textbf{valuation} $v(\mathcal{X})$. 
\end{frame}
} \else {
\begin{frame}{효용 모형}
$c_j$에 다니면 $t_j \geq 0$ 단위의 \textbf{효용}이 발생한다. Wlog, $t_j \leq t_{j+1}$.

@@

\ifen {
\begin{frame}[plain]
\begin{center}
~
\begin{beamercolorbox}[wd=.5\textwidth,sep=8pt,center,shadow=false,rounded=true]{section in head/foot}
    \usebeamerfont{title} \insertsectionhead
  \end{beamercolorbox}
\end{center}
\end{frame}
} \else {
\begin{frame}[plain]
\begin{center}
~
\begin{beamercolorbox}[wd=.5\textwidth,sep=8pt,center,shadow=false,rounded=true]{section in head/foot}
    \usebeamerfont{title} \insertsectionhead
  \end{beamercolorbox}
\end{center}
\end{frame}
} \fi

\begin{frame}[plain]%{Safety, target, and reach schools}

@@

\ifen {
\begin{frame}[plain]
\begin{center}
~
\begin{beamercolorbox}[wd=.5\textwidth,sep=8pt,center,shadow=false,rounded=true]{section in head/foot}
    \usebeamerfont{title} \insertsectionhead
  \end{beamercolorbox}
\end{center}
\end{frame}
} \else {
\begin{frame}[plain]
\begin{center}
~
\begin{beamercolorbox}[wd=.5\textwidth,sep=8pt,center,shadow=false,rounded=true]{section in head/foot}
    \usebeamerfont{title} \insertsectionhead
  \end{beamercolorbox}
\end{center}
\end{frame}
} \fi

@@

%\looseness=-1
Using a variable-elimination technique, we reduce the cost of computing $v(\mathcal{X})$ to amortized unit time, obtaining an $O(hm)$ algorithm. Our Julia implementation (Kapur 2022) solves an $m = 16384$ instance in \rlap{200 ms.}

@@

\ifen {
\begin{frame}[plain]
\begin{center}
~
\begin{beamercolorbox}[wd=.5\textwidth,sep=8pt,center,shadow=false,rounded=true]{section in head/foot}
    \usebeamerfont{title} \insertsectionhead
  \end{beamercolorbox}
\end{center}
\end{frame}
} \else {
\begin{frame}[plain]
  \vfill
\begin{center}
~
\begin{beamercolorbox}[wd=.5\textwidth,sep=8pt,center,shadow=false,rounded=true]{section in head/foot}
    \usebeamerfont{title} \insertsectionhead
  \end{beamercolorbox}
\end{center}
\end{frame}
} \fi

@@

Kapur, Max and Sung-Pil Hong. 2022. ``The College Application Problem.'' Preprint. \url{https://arxiv.org/abs/2205.01869}.

@@
